### 2018-06-03 MSc_Marko_Docevski_tekst.tex
Updates from ShareLaTeX

\subsection{Граматички / експертски системи}

@@

\subsection{Алгоритми со длабоко учење}

@@

\cite{Eck2002} Pure RNNs are not very well suited for music sequence learning due to the problem of the vanishing/explodin gradient. In music longer term dependencies is what defines the syle or piece, events spanning from several notes to many, many measures. CONCERT and other previuos works employed music representation with a psychologically-realistic distributed input encoding, that gave it a bias towards chromatically and harmonically related notes. By representing in a single timestep a note rather than a slice of time, the number of time steps to be bridge by the network in learning global structure is greatly reduced.They prefered a localist representation of the notes, but with a straightforward encoding. 0/1 values for notes being off/on. Make no distinction between melody/chords. They used a slice of time representation. Quantization is the size of the smallest note size in the datase 1/8. Supposedly this is better for LSTM because it'd force the network to learn relative duratoins of notes. No explicit way to denote note endings. 1/4 note is represented the same as 2 1/8 of the same pitch. 2 ways to deal with this, have the quantization 1/2 of the size of the smallest note size and mark note endings with a zero, or have special markings for note starts, but unsure how this would scale in multi-voice melodies. Very short, fixed sized samples, 12 bar blues, 8 notes per bar. 2 experiments: 1) only chord progression, 2) chord progression + melody line. Experiment 1) 4 cell blocks containing 2 cells each, fully connected to each orher and to the input layer. Output layer fully connected to all cells and to the input layer. Experiments showed that learning was faster after the network made a big mistake or several small ones. Activation of output logistic sigmoid. Network was trained with cross-entropy as the objective function, RMSE considered inadecvate .Chord notes were predicted using a decision threshold of 0.5. Experiment 2) network topology, 8 cell blocks of 2 cells each, four of the cell blocks are fully connected to the input units for chods. The other four cell blocks are fully connected to the input units for melody. The chord cell blocks have recurrent connecctions to themselves and to the melody cell blocks. The melody cell blocks are only recurrently connected to melody cell blocks. Melody information does not reach the cell blocks responsible for generation chords.In the output layer, units for chords are f.c. to c.bs. for chords and i.u. for chords.O.us for melody are f.c. to c.bs for melody and i.u. for melody. Melody output activations are adjusted so that they sum up to 1.0. Then using a uniform random number in the range [0,1] the appropriate next note is chosen. Music composed by the system is considered good by the author, but the problem it solves is very contained and strictly defined. Still no objective anaysis has been conducted. There was no variety in the chod structure, so it's best understood as improvisation over a predefined form rather than composition. The divide between the melody and chord sections of the networks, and the chords influencing the melody but not vice versa is believed to be analogus to reality, where the rhythm section influences the solo is but not the other way around.

\cite{Eck2008} Long-timescale structure is fundamental to music, so all architectures try to capture it. However regardles of the architecture the possibilities explode with the timescale. Their solution primarily deals with sequences extracted from MIDI, not performed music. The core algorithms are expected to be well suited for digital audio as well. Improvement over the previous work is addition of time-delay connections that correspond to the metrical hierarchy of a particular piece. This meter information should provide the LSTM net with musically important temporal structure. With it the model was able to capture some of the repetitive structure crucial to learning a musical style. Meter is the sense of strong and weak beats that arises from the interaction among hierarchical levels of sequences having nested periodic components. Such a hierarchy is implied in Westernmusic notation, where different levels are indicated by kinds of notes (whole notes, half notes, quarter notes, etc.) and where bars establish measures of an equal number of beats (Handel, 1993). For instance, most contemporary pop songs are built on four-beat meters. In such songs, the first and third beats are usually emphasized. Knowing the meter of a piece of music helps in predicting other components of musical structure such as the location of chord changes and repetition boundaries (Cooper and Meyer, 1960). Meter provides us with key information about musical structure. Music, at least popular Western music, tends to be chunked in ways that correspond to meter. Chord changes, for example, usually occur on metrical boundaries. Also, music tends to repeat at intervals corresponding to the metrical hierarchy. A dynamic learner such as a recurrent neural network (details described below) could learn to repeat a fixed number of learned patterns, but it could not learn to repeat an arbitrary sequence because it has no way to implement content-addressable memory. Meter is provided to the network in the form of delayed inputs. If music is sampled k times a measure, eg k=8, for 4/4 meter to make it easy for the network to correlate metrically-relevant delays by providing time-delayed copises of the input at times t -15, t-31 and t-63 (2,4 and 8 measures). The recurrent network will learn the regular sequence on its own, but lags related to meter are give more attention. Theiir LSTM network has a standard feed-forward layer in paralel to the LSTM cells. It should accelerate training by learning local dependencies quicker than LSTM alone, leaving the LSTM cells to handle longer timescale dependencies. This reulted in melodies with smoother contours. The network is trained as a next-step predictor. It trains to predict probability density over all possible notes at time t using as input the note (and chord) values at time t-1. In this model the network recieves as input not just the sequence delayed by a single lag (t-1) but also delayed by lags correpsonding to metrical structure of the piece. Multiple songs are presented to the network in a single sequence. Error flow is truncated at song boundaries, so that the network doesn't learn spurious corelations from one song to another. With the trained model, music can be generated by presenting it with the first few notes of a song that it has never seen in training and using the network predictions to generate network inputs. Network predictions are conditioned using a softmax function, ensuring that the sum of the output vector is 1.0. This allows us to interpret the output vector as a probability estimation from which we can select the next note. The selected note is then presented to the network at the next timestep as an input. For some simulations, we applied a threshold to our note generation, ensuring that very low probability notes would not be chosen. The threshold we used was 1/N where N is the cardinality of the output vector. MIDI data representation, turned into input vectors by sampling the file at eight-note intervals. Most other data formats can be turned into MIDI files. Number of octaves limited between C3 and C5, notes outside of range folded into range. Time is represented implicitly, so no note durations are present in i/o. Identical to previoys works. Notes are represented locally in a 1-hot vector. Notes not in the corpus are not represented. Chords are also encoded locally in 1-hot vector. So eg. Fmaj7 is represented by only 1 input element and not as several units representing the notes that make up the chord. Advantage of this aproach is the network can more quickly learn chord structure, but distadvantage is the network cannot generalize to unseen chords. Several 1-hot vectors are added to encode metrical structure, corressponding to the time-delayed versions of the input. Dataset: They used Irish reels in 4/4. They selected reels from thesession.org that were in C maj and E maj. It's a good dataset with at least 1700 reels. It however does not have consistant chord information on all songs, so they used only the melodies. They also used the Nottingham dataset, transposing all the songs in the same key. It does provide good chord information that they used. THe most interesting experiment, run on Nottingham datase, contains both melody and chords. Take care not to overfit the dataset with too hight a network capacity (too many nodes) or too much training, it'll just repeat the same notes. When training on melodies and chrods, they achieved a result where the melody does follow the chords. Also the model did capture the slow-changing chord structures and is able to synchronize faster-changing melodic structure with the chords.

>>>>

\section{Граматички / експертски системи}

@@

\ection{Алгоритми со длабоко учење}

@@

Првиот документиран обид за користење на невронски мрежи за генерирање на музика е од страна на Д. Ек во трудот \cite{Eck2002}. Целта е тренирање на модел кој ќе генерира блуз во 12 такта. Музиката се состои од 2 оделни компоненти, секвенца на акорди кои го одредуваат ритамот и ја даваат основната структура на музиката и мелодиска линија. Моделот има две групи од по 4 пара LSTM (Long Short Term Memory) ќелии (рекурентни неврони), од кои едната е задолжена за учење на мелодијата, а друга за акордната секвенца. Музиката ја преставуваат во тн. репрезентација во пиано лента, која е дво димензионална бинарна матрица во која едната оска (подолгата) го претставува времето, а другата оска го преставува множеството од можни тонови. Музиката има строга структура, 12 такта, 4/4 ритам и времето е подделено на 1/8 ноти. Моделот го испробале на два експерименти, во првиот тој ја учи само секвенцата на акорди, а во вториот и акордите и мелодијата. Тренирањето е извршено со cross-entropy фунција на загува, со сигмоидална функција на активација. Делот за учење на акорди ја споделува својата внатрешна состојба со делот за учење на мелодијата. Авторот смета дека системот генерира добра музика, но сепак проблемот што пробува да го реши е хипер фокусиран и ограничен. Трудот \cite{Eck2008} претставува надоградување на претходно опишаниот систем. Го нема веќе фокусот на блуз во 12 такти, туку работат со податочно множество од поразлини MIDI датотеки, зголемена е комплексноста на моделот, многу повеќе ќелии, повеќе слоеви од истите. Додатно

Long-timescale structure is fundamental to music, so all architectures try to capture it. However regardles of the architecture the possibilities explode with the timescale. Their solution primarily deals with sequences extracted from MIDI, not performed music. The core algorithms are expected to be well suited for digital audio as well. Improvement over the previous work is addition of time-delay connections that correspond to the metrical hierarchy of a particular piece. This meter information should provide the LSTM net with musically important temporal structure. With it the model was able to capture some of the repetitive structure crucial to learning a musical style. Meter is the sense of strong and weak beats that arises from the interaction among hierarchical levels of sequences having nested periodic components. Such a hierarchy is implied in Westernmusic notation, where different levels are indicated by kinds of notes (whole notes, half notes, quarter notes, etc.) and where bars establish measures of an equal number of beats (Handel, 1993). For instance, most contemporary pop songs are built on four-beat meters. In such songs, the first and third beats are usually emphasized. Knowing the meter of a piece of music helps in predicting other components of musical structure such as the location of chord changes and repetition boundaries (Cooper and Meyer, 1960). Meter provides us with key information about musical structure. Music, at least popular Western music, tends to be chunked in ways that correspond to meter. Chord changes, for example, usually occur on metrical boundaries. Also, music tends to repeat at intervals corresponding to the metrical hierarchy. A dynamic learner such as a recurrent neural network (details described below) could learn to repeat a fixed number of learned patterns, but it could not learn to repeat an arbitrary sequence because it has no way to implement content-addressable memory. Meter is provided to the network in the form of delayed inputs. If music is sampled k times a measure, eg k=8, for 4/4 meter to make it easy for the network to correlate metrically-relevant delays by providing time-delayed copises of the input at times t -15, t-31 and t-63 (2,4 and 8 measures). The recurrent network will learn the regular sequence on its own, but lags related to meter are give more attention. Theiir LSTM network has a standard feed-forward layer in paralel to the LSTM cells. It should accelerate training by learning local dependencies quicker than LSTM alone, leaving the LSTM cells to handle longer timescale dependencies. This reulted in melodies with smoother contours. The network is trained as a next-step predictor. It trains to predict probability density over all possible notes at time t using as input the note (and chord) values at time t-1. In this model the network recieves as input not just the sequence delayed by a single lag (t-1) but also delayed by lags correpsonding to metrical structure of the piece. Multiple songs are presented to the network in a single sequence. Error flow is truncated at song boundaries, so that the network doesn't learn spurious corelations from one song to another. With the trained model, music can be generated by presenting it with the first few notes of a song that it has never seen in training and using the network predictions to generate network inputs. Network predictions are conditioned using a softmax function, ensuring that the sum of the output vector is 1.0. This allows us to interpret the output vector as a probability estimation from which we can select the next note. The selected note is then presented to the network at the next timestep as an input. For some simulations, we applied a threshold to our note generation, ensuring that very low probability notes would not be chosen. The threshold we used was 1/N where N is the cardinality of the output vector. MIDI data representation, turned into input vectors by sampling the file at eight-note intervals. Most other data formats can be turned into MIDI files. Number of octaves limited between C3 and C5, notes outside of range folded into range. Time is represented implicitly, so no note durations are present in i/o. Identical to previoys works. Notes are represented locally in a 1-hot vector. Notes not in the corpus are not represented. Chords are also encoded locally in 1-hot vector. So eg. Fmaj7 is represented by only 1 input element and not as several units representing the notes that make up the chord. Advantage of this aproach is the network can more quickly learn chord structure, but distadvantage is the network cannot generalize to unseen chords. Several 1-hot vectors are added to encode metrical structure, corressponding to the time-delayed versions of the input. Dataset: They used Irish reels in 4/4. They selected reels from thesession.org that were in C maj and E maj. It's a good dataset with at least 1700 reels. It however does not have consistant chord information on all songs, so they used only the melodies. They also used the Nottingham dataset, transposing all the songs in the same key. It does provide good chord information that they used. THe most interesting experiment, run on Nottingham datase, contains both melody and chords. Take care not to overfit the dataset with too hight a network capacity (too many nodes) or too much training, it'll just repeat the same notes. When training on melodies and chrods, they achieved a result where the melody does follow the chords. Also the model did capture the slow-changing chord structures and is able to synchronize faster-changing melodic structure with the chords.
